Optimal management of a stochastically varying population when policy adjustment is costly

### 2014-12-19 manuscripts/manuscript.Rmd
update Dockerfile, tweak manuscript, separate out supplement

```{r configure, include=FALSE, cache=FALSE}
#source("components/install.R")

library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)
# PDF-based figures
opts_chunk$set(dev='pdf')
fig.cache <- TRUE

@@

```{r Figure_4, fig.cap="Variance and autocorrelation of harvest dynamics, as a function of the penalty coefficient.  Adjustment penalty is calculated as percentage of the maximum expected $NPV$ in the basic case with no policy adjustment costs.", fig.height=8, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png")}

@@

```{r}
fig5_caption <- paste("Distribution of profits from fishing and costs paid to adjustment, under linear costs ($\Pi_1$). (Alternate penalties can be found in the supplementary materials)  Distributions are shown when managing a system in which the adjustment costs present compared to the same costs and profits of managing the system using a Reed model, which ignores adjustment costs, and thus tends to pay higher fees for adjustment.  In contrast, the difference in profit after accounting for costs is only marginally less than the value derived by managing the population to optimize profits when ignoring adjustment costs (Reed model).")
```

@@

Estimating policy adjustment costs and how they respond to the size of
proposed policy changes would be empirically challenging. Recognizing
this fact, in our analyses, we compared different plausible forms that
these costs might take. We compared two cost structures where we assumed
the magnitude of policy adjustment costs increased with the magnitude
of the change in policy being proposed with an alternative formulation
in which we assumed there was a fixed cost associated with making any
change to current policy. While each formulation clearly provides only a
very phenomenological representation of the policy setting environment,
we believe that each enables us to explore meaningful differences in how
policy adjustment costs might operate. That being said, in the real-world
we might expect the different types of adjustment cost to operate in
combination.

The biggest differences between the representations of policy adjustment
cost that we consider are between those that have a smoothing effect
on annual quotas and those that do not. The few past studies that
incorporate policy adjustment costs in models of fisheries management
[@Ludwig1980; @Feichtinger1994; @Wirl1999] assumed costs of
policy adjustment increased as a quadratic function of the magnitude of
the policy change being proposed (our $\Pi_2$ formulation) In
effect, this means that small changes to annual quotas incur little
extra cost, but large changes to annual quotas become disproportionately
expensive to make. Including costs of this form smoothes inter-annual
variation in the catch quotas that would be recommended. Also, the catches and remaining stock sizes
corresponding to optimal management become more autocorrelated in time,
because it takes the fishery longer to harvest down peaks in abundance
that follow large recruitment pulses. The effects of smoothing here are
similar to those predicted when assuming the cost per unit effort is
increasing in the amount of effort expended (e.g., @Lewis1981, @McGough2009; 
see Supporting Information for a summary of the relevant
results) as opposed to associating extra costs with changes to policy
*per se*. Smoothing effects of this type are what one might have
expected when including policy adjustment costs.

@@

That the different representations of policy adjustment costs result in
such different dynamics suggests that researchers constructing fisheries
economics models should proceed cautiously when choosing how to represent
these costs. However, in our own experience, we have found that while
modeling studies sometimes mention policy adjustment costs when motivating
model assumptions (references), they rarely provide much justification
for the choice of functional form used or test the sensitivity of any
conclusions drawn to alternative specifications. Indeed, we were surprised
to find such clear differences between the functional forms, because
we anticipated that the limiting case of increasing policy adjustment
costs within each functional form should be the same, namely a constant
annual quota that does not change through time (open-loop control).

@@

Assuming an adjustment cost exists when in fact it is absent is not the
same as using the wrong functional form when it is present.  Here, our
results showed that choice of the functional form makes a significant
difference.  We show it can be better to ignore policy costs than to derive a 
policy based on assuming the wrong functional form.  This result further
underscores the importance of modelers using caution when seeking to account
for these costs. Assuming an arbitrary form in order to capture the influence
of adjustment costs is thus unlikely to be instructive.

@@

A worthwhile modeling extension suggested by our results would be to
examine the implications of policy adjustment costs for risks of stock
collapse. Examining risks of stock collapse is not possible with the
Reed model-formulation that we followed here but would be worthwhile
in light of our finding that the variance and autocorrelation in stock
sizes through time can be affected when accounting for costs of policy
adjustment. Both are properties known in other modeling settings to
be associated with changes to the risk of extinction or of transitions
between alternative stable states [@Scheffer2009; @Boettiger2013].

<!-- Comment on increasing variance or increasing autocorrelation from changes in policy costs / optimal management-->

@@

# Supplementary material #

##  Impact on the stock dynamics ##

```{r StockDynamics, dependson="tidy", fig.cap="Example realization of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=5, fig.width=8, cache = fig.cache, dev=c("pdf", "svg", "png")}

dt_s <- dt %>% 
  filter(replicate=="rep_17" & time < 20) %>%
  select(time, fishstock, alternate, penalty_fn) %>%
  gather(variable, value, -time, -penalty_fn)

ggplot(dt_s, aes(time, value, col=variable, lty=variable)) +
  geom_line(lwd=1) +
  facet_grid(penalty_fn~., labeller = labeller) + 
  labs(x="time", y="stock size", title = "Example Stock Dynamics")  +
  scale_color_discrete(labels=c("With Penalty", "Reed Solution"), name="") +
  scale_linetype_discrete(labels=c("With Penalty", "Reed Solution"), name="")

```{r StockStats, fig.cap="Variance and autocorrelation of stock dynamics, as a function of the penalty coefficient.  Adjustment penalty is calculated as percentage of the maximum expected $NPV$ in the basic case with no policy adjustment costs.", fig.height=8, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png")}

stock_stats <- filter(figure4_df, timeseries=='stock')
ggplot(stock_stats, aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn)) + 
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  coord_cartesian(xlim = c(0, .3)) + 
  facet_wrap(~ statistic, scale="free_y") + 
  scale_color_discrete(labels=relabel, name="Penalty function") + 
  scale_fill_discrete(labels=relabel, name="Penalty function") +
  xlab("Penalty size (as fraction of NPV0)")

Recall that in our model definition, stock sizes are measured before harvest ($h_t$ being determined after observing stock size
$x_t$) and thus show less influence of the policy cost than the harvest itself.

## Additional stochastic realizations ##

Figure 3 of the main text shows a single stochastic realization.  Here we provide a sample of further realizations (including both stock and harvest dynamics).  Full

```{r Figure_S2, dependson="tidy", fig.cap="Multiple realizations of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=10, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png"), cache=FALSE}

labels <- c(harvest = substitute(paste("Harvest Policy under ", paste(Pi[1-3]))), 
            harvest_alt = substitute(paste("Harvest under ", paste(Pi[0]), " (Reed)")),
            fishstock = substitute(paste("Stock dynamics under under ", paste(Pi[1-3]))), 
            alternate = substitute(paste("Stock under ", paste(Pi[0]), " (Reed)")))
labeller <- function(variable,value){
if (variable=='penalty_fn') {
    return(relabel[paste(value)])
  } else {
    return(value)
  }
}

ex <- dt %>% 
  filter(replicate %in% paste0("rep_", 16:20) & time < 20) %>%
  select(time, fishstock, harvest, alternate, harvest_alt, replicate, penalty_fn) %>%
  gather(variable, stock, -time, -replicate, -penalty_fn)

ggplot(ex) + 
  geom_line(aes(time, stock, col=variable), lwd=1) + 
  scale_color_discrete(labels=labels, name="") +  
  facet_grid(replicate ~ penalty_fn, labeller=labeller)

write.csv(ex, "components/data/figure_s2.csv")
```

## Economic penalties using quadratic costs ##

```{r quadraticcosts}

source("components/quad_costs.R")
```

```{r Figure_S1, dev=c('pdf', 'svg'), cache=FALSE}

labels <- c(harvest = substitute(paste("Harvest Policy under ", paste(Pi[4]))), 
                                harvest_alt = substitute(paste("Harvest under ", paste(Pi[0]), " (Reed)")))

ex <- dt_quad %>% 
  filter(replicate=="rep_17" & time < 20) %>%
  select(time, harvest, harvest_alt) %>%
  gather(variable, stock, -time) 
ggplot(ex) + 
  geom_line(aes(time, stock, col=variable), lwd=1) + 
  scale_color_discrete(labels=labels, 
                       name="")

```{r quadshiftrate}
shifts <- dt[, fraction_no_shift(harvest), by=c("replicate")]
ave <- shifts[, mean(V1)]
```

Like the $\Pi_2$ adjustment cost based penalty, this quadratic cost based penalty shows periods of contant, strictly positive policy only `r ave * 100`% of the time, and exhibits the same kind of smoothing of more small adjustments that is seen in $\Pi_2$ scenario.

<!-- Figure 5 extended across -->

```{r Figure_S3, fig.cap="Distribution of profits from fishing and costs paid to adjustment.  Distributions are shown when managing a system in which the adjustment costs present are determined by the model indicated, and compared to the same costs and profits of managing the same system using a Reed model, which ignores adjustment costs, and thus tends to pay much higher fees for adjustment.  In contrast, the difference in profit after accounting for costs is only marginally less than the value derived by managing the population to optimize profits when ignoring adjustment costs (Reed model).", fig.width=8, fig.height=6, dependson="histogram-calc", dev=c("pdf", "svg", "png"), cache=FALSE}

assume <- levels(as.factor(hist_dat$Assumption))
labeller <- function(variable,value){
if (variable=='penalty_fn') {
    return(relabel[paste(value)])
  } else {
    return(paste(value))
    #return(assume[paste(value)])
  }
}

ggplot(hist_dat) + 
  geom_density(aes(value, fill=variable, color=variable), alpha=0.8)+
  facet_grid(Assumption~penalty_fn, labeller = labeller)

```{r Figure_S4, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue), shown by penalty function over increasing magnitudes of cost.  All values are relative to the cost-free adjustment optimum, NPV0.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. The severity of the latter mistake increases sharply as adjustment costs make a larger fraction of the total costs.", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}

ggplot(table1_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, ncol=2) + 
  scale_x_discrete(labels = c('L1' = substitute(paste(Pi[1])),
                              'L2' = substitute(paste(Pi[2])),
                              'fixed' = substitute(paste(Pi[3])))) + 
  xlab("Penalty function")
write.csv(table1_long, "components/data/figure6.csv")
```

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure S5 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10% (panel A)
less than optimal free based on the $c_i$ calibration. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

For fixed costs ($\Pi_3$), the reverse is true.  Because fixed costs increase the
volatility of the optimal solution while $\Pi_2$ costs decrease it, simply
ignoring the cost of adjustment all together results in a higher value
than either assuming costs are quadratic ($\Pi_2$) when they are fixed ($\Pi_3$),
or assuming they are fixed when they are quadratic.  Scenarios in which true costs are quadratic (regardless of what
cost structure is assumed by the policy) do significantly worse than
those that are not (i.e. all cases that assume quadratic but actually
have linear or fixed costs).

These patterns hold across the different size reductions shown (from 10%
to 25%, panels a-d respectively), though net present value becomes negative under the quadratic
form for the larger penalties as costs paid for adjustments exceed
profits derived from harvest.  (Though optimal solutions will always
avoid negative net present values, these scenarios in which policies are
applied under conditions other than assumed in the optimality calculation
have no such guarantee.) For each scenario, results are averaged over
100 replicates.

```{r Figure_S5, fig.cap="Mismatches. X axis labels indicate the model used to derive the policy, followed by the model used to simulate the penalty. Blue bars show the net present value of that scenario.  Compare to red bars, which show the consequence of using a policy that ignores costs all together in the same scenario (i.e. costs as indicated by second code), rather than assuming the policy indicated by the first code in the pair.  All values are scaled to a fraction of the adjustment free net present value.", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

ggplot(filter(table2_long, reduction==0.25), 
  aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
write.csv(table2_long, "components/data/figure7.csv")
```

<!--

```{r Figure_S5, fig.cap="Effect of varying severity of adjustment costs on mismatches. Successive panels show increasing reduction of the net present value by increasing penalty coefficients: e.g. a policy in the first panel would lose only 10% (panel A) of its net present value to adjustment costs if it had the correct form of those costs, as determined by the choice of penalty coefficient.  Plots structured as in figure 7, which corresponds to the fourth panel", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}
ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
write.csv(table2_long, "components/data/figure7.csv")
```

-->

>>>>

```{r configure, include=FALSE, cache=FALSE}
library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)
# PDF-based figures
opts_chunk$set(dev='pdf')
fig.cache <- TRUE

@@

```{r Figure_4, fig.cap="Variance and autocorrelation of harvest dynamics, as a function of the penalty coefficient.  Adjustment penalty is calculated as percentage of the maximum expected $NPV$ in the basic case with no policy adjustment costs.", fig.height=4, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png")}

@@

```{r}
fig5_caption <- paste("Distribution of profits from fishing and costs paid to adjustment, under linear costs ($\\Pi_1$). (Alternate penalties can be found in the supplementary materials)  Distributions are shown when managing a system in which the adjustment costs present compared to the same costs and profits of managing the system using a Reed model, which ignores adjustment costs, and thus tends to pay higher fees for adjustment.  In contrast, the difference in profit after accounting for costs is only marginally less than the value derived by managing the population to optimize profits when ignoring adjustment costs (Reed model).")
```

@@

Estimating policy adjustment costs and how they respond to the size of
proposed policy changes would be empirically challenging. Recognizing
this fact, we compared different plausible forms that
these costs might take. We compared two cost structures where we assumed
the magnitude of policy adjustment costs increased with the magnitude
of the change in policy being proposed with an alternative formulation
in which we assumed there was a fixed cost associated with making any
change to current policy. While each formulation clearly provides only a
very phenomenological representation of the policy setting environment,
we believe that each enables us to explore meaningful differences in how
policy adjustment costs might operate. That being said, in the real-world
we might expect the different types of adjustment cost to operate in
combination.

The biggest differences between the representations of policy adjustment
cost that we consider are between those that have a smoothing effect
on annual quotas and those that do not. The few past studies that
incorporate policy adjustment costs in models of fisheries or other environmental management contexts [@Feichtinger1994; @Wirl1999] have assumed costs of
policy adjustment increased as a quadratic function of the magnitude of
the policy change being proposed (analogous to $\Pi_2$ formulation). In
effect, this means that small changes to annual quotas incur little
extra cost, but large changes to annual quotas become disproportionately
expensive to make. Including costs of this form smooths inter-annual
variation in the catch quotas that would be recommended. Also, the catches and remaining stock sizes
corresponding to optimal management become more autocorrelated in time,
because it takes the fishery longer to harvest down peaks in abundance
that follow large recruitment pulses. The effects of smoothing here are
similar to those predicted when assuming the cost per unit effort is
increasing in the amount of effort expended (e.g., @Lewis1981, @McGough2009; 
see Supporting Information for a summary of the relevant
results) as opposed to associating extra costs with changes to policy
*per se*. Smoothing effects of this type are what one might have
expected when including policy adjustment costs.

@@

That the different representations of policy adjustment costs result in
such different dynamics suggests that researchers constructing fisheries
economics models should proceed cautiously when choosing how to represent
these costs. However, in our own experience, we have found that while
modeling studies sometimes mention policy adjustment costs when motivating
model assumptions [e.g. @Feichtinger1994; @Wirl1999; @Lewis1981; @McGough2009], they rarely provide much justification
for the choice of functional form used or test the sensitivity of any
conclusions drawn to alternative specifications. Indeed, we were surprised
to find such clear differences between the functional forms, because
we anticipated that the limiting case of increasing policy adjustment
costs within each functional form should be the same, namely a constant
annual quota that does not change through time (open-loop control).

@@

Assuming an adjustment cost exists when in fact it is absent is not the
same as using the wrong functional form when it is present.  In the supplement we show that choice of the functional form makes a significant
difference.  We show it can be better to ignore policy costs than to derive a 
policy based on assuming the wrong functional form.  This result further
underscores the importance of modelers using caution when seeking to account
for these costs. Assuming an arbitrary form in order to capture the influence
of adjustment costs is thus unlikely to be instructive.

@@

A worthwhile modeling extension suggested by our results would be to
examine the implications of policy adjustment costs for risks of stock
collapse. Examining risks of stock collapse is not possible with the
Reed model-formulation that we followed here but would be worthwhile
in light of our finding that the variance and autocorrelation in stock
sizes through time can be affected when accounting for costs of policy
adjustment. Both are properties known in other modeling settings to
be associated with changes to the risk of extinction or of transitions
between alternative stable states [@Scheffer2009]. This highlights the risk
that changes in management policies can either mimic or mask such possible
early warning signs of sudden transitions, and points to another way
in which more context is needed before such approaches can be useful to
management [@Boettiger2013].

@@
